Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-15 chapters/LBM_technique.tex
Addressed some typesetting and more notational issues in methods section

\begin{equation}
    \sigma = -p\mathbb{I} + \eta\left( \nabla \vec{u} + (\nabla \vec{u})^T -\frac{2}{3}(\nabla \vec{u})\mathbb{I} \right)
\end{equation}

Substituting this stress tensor into the momentum equation and using the relation, $\nabla\cdot \vec{u} = 0$ from the incompressibility constraint, the incompressible 
Navier-Stokes equation can be derived which in vector notation is given by,

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = -\nabla p + \eta \nabla^2 \vec{u} + \vec{f}
\end{equation}

To facilitate numerical simulation and analysis, it is often useful to express the Navier-Stokes equation in non-dimensional form. Introducing characteristic scales for length 
$l_0$, time $t_0$, and density $\rho_0$, dimensionless variables are defined as

\begin{equation}
    \bar{l} = \frac{l}{l_0}, \quad \bar{t} = \frac{t}{t_0}, \quad \bar{\rho} = \frac{\rho}{\rho_0}, \quad \bar{\vec{u}} = \frac{\bar{l}}{\bar{t}}, 
\end{equation}

Substituting these into the incompressible Navier-Stokes equation leads to the dimensionless form:

\begin{equation}
    \frac{\partial \bar{\vec{u}}}{\partial \bar{t}} + (\bar{\vec{u}} \cdot\nabla)\bar{\vec{u}} = \frac{1}{Re} \nabla^2 \bar{p} = \frac{p}{\rho \vec{u}^2} - \nabla\bar{p}
\end{equation}

where the Reynolds number is defined as $Re = \frac{\rho_0 u_0 l_0}{\eta}$ which characterizes the ratio of inertial to viscous forces in the system.

@@

Bijels form through spinodal decomposition of two partially miscible fluids, a process that can be effectively modeled using the Cahn-Hilliard equation. \cite{cahn_spinodal_1961} While the
original equation is written for a concentration of species $c$. For a binary mixture, it can be recast in terms of a conserved scalar order parameter $\phi$, typically representing the 
local concentration difference between the two fluid components. As we are also accounting for the effect of hydrodynamics on the flow of mass, the advection term
from the mass flux $\vec{J}$ cannot be set to 0. Therefore the Cahn-Hilliard equation is given by,

\begin{equation}
    \begin{split}
        \nabla \cdot \vec{J} &= \nabla \cdot \vec{u} - M \nabla \mu \\
        \frac{\partial \phi}{\partial t} + \nabla \cdot \vec{J} &= 0 \\
        \frac{\partial \phi}{\partial t} + \nabla \cdot \phi \vec{u} &= \nabla \cdot \left( M \nabla \mu \right)
    \end{split}
\end{equation}

Here, $M$ is the mobility coefficient and $\mu$ is the chemical potential, which is derived from the variation of the free energy functional $F[\phi]$ which is the sum of a bulk free energy
$f(\phi)$ and an interfacial contribution. This formulation allows for the modeling of interfacial tension and the spontaneous formation of bicontinuous domains characteristic of spinodal decomposition.
To capture not only the diffusive separation of phases but also the resulting fluid motion, the Cahn-Hilliard equation is coupled with the Navier-Stokes equation. This is done
through the pressure which is expressed as the sum of the mechanical $p$ and chemical pressure $p_{chem}$ contributions to yield $p_{tot} = p + p_{chem}$. For a binary system, $p_{chem}$ 
can be written as a function of $\phi$ given by,

\begin{equation}
    p_{chem} = \left[ \phi \mu - f(\phi) - \kappa(\phi \nabla^2 \phi + \frac{1}{2}|\nabla \phi |^2 )\right] \mathbb{I} + \kappa \nabla \phi \otimes \nabla \phi
\end{equation}

The divergence of the above expression is found to be $\nabla \cdot p_{chem} = -\phi \nabla \mu$. Substituting the $p_{tot}$ for $p$ in the Navier Stokes equation,
the Cahn-Hilliard-Navier-Stokes (CHNS) framework is obtained. This coupling is essential in systems like bijels, where advection play a significant role in domain 
coarsening and morphology arrest. The momentum conservation equation in this framework is given by,

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = \eta \nabla^2 \vec{u} - \nabla p - \phi \nabla \mu
\end{equation}

@@

However the BGK operator controls all moments at a single relaxation rate which can induce numerical instabilities and inaccuracies
at higher fluid velocities. \cite{liu_simulation_2023, adhikari_fluctuating_2005} These numerical instabilities can be resolved by using
a Multiple Relaxation Time (MRT) collision operator. However the relaxation rates used in an MRT collision operator must be tuned carefully
to ensure that the appropriate fluid properties are modelled. Simpler variants of the MRT collision operator such as the Two Relaxation Time (TRT)
collision operator have been proposed as a solution that provides some of the benefits of MRT models with the simplicity of SRT models. These 
have facilitated simulation of a wider range of physical phenomena \cite{adhikari_fluctuating_2005, liu_simulation_2023}. The SRT collision 
operator is selected due to the additional computational performance needed by MRT not being needed. The combined streaming and 
collision step in LBM, using the BGK operator, is expressed in Equation \ref{eq:LBM_BGK}.

@@

$\rho$ and $\textbf{u}$ in Equation \ref{eq:LBM_Feq} are defined as the macroscopic parameters for density and velocity 
and can be calculated from the mass distribution using $\rho = \sum f_i$ and $\rho \vec{u} = \sum f_i \vec{c}_i$ 
respectively. Using the Chapman-Enskog expansion, the Navier-Stokes equation at the incompressible limit at low Mach
numbers can be recovered. \cite{qian_lattice_1992, he_lattice_1997} The regime accessible through this method is suitable 
for simulations in this work as the $ 1 \leq Re \leq 100 $ and the mach number, $Ma < 0.01$, fulfilling the stability 
criterion and usage requirements for the presented hydrodynamic model.

The density and velocity of the fluid at each grid point is calculated from the zeroth and first moment of the particle density
distribution,

@@

The momentum flux tensor is calculated from the Chapman-Enskog expansion of \ref{eq:LBM_BGK} to the second order yielding,

@@

The momentum flux tensor can be further expressed as a sum of the equilibrium and non-equilibrium components of the stress tensor, denoted by 0 and 1 given by,

@@

An effective mass of each fluid at node $\vec{x}$ is used in place of the actual density to scale it between zero 
and one and is defined as $\psi^{k}(\vec{x},t) = \rho_{0}\left[1 - \exp(-\frac{\rho^{k}(\vec{x}, t)}{\rho_{0}})\right]$. 
In this model, the SC force is incorporated into the macroscopic velocities that are then used to calculate the equilibrium
distribution $f_{i}^{k, eq}$ for fluid $k$, defined as,

@@

The particles are discretized on the lattice according to the method laid out in Ladd 1994 and Aidun 1998
\cite{ladd_numerical_1994,aidun_direct_1998}. Nodes representing the particle are marked as solid nodes that replicate
a no-slip boundary condition through a moving bounce-back boundary. This is implemented into the distribution function
by reflecting the outgoing populations of $f_i^k$ to the opposite lattice velocity

@@

This facilitates momentum exchange between particle and fluid which can be calculated analytically as 
\(\Delta\vec{p}^k_i \frac{\Delta t}{(\Delta x)^3} = 2 f^{k,\star}_i(\vec{x},t)\vec{c}_i - \frac{2w_i}{c_s^2}\rho(\vec{u}_i\cdot\vec{c}_i)\vec{c}_i\).
The sum of the momentum change across the surface of the particle is computed to obtain the force and torque on the particle,

@@

@@

where $R_i$ and $R_j$ are the radii of each particle involved in the interaction, $\vec{r}_{ij}$ is the distance
vector between the particle centers, $\vec{u}_{ij}$ are the relative velocities of the particles and $\Delta_c$ 
is the cutoff distance when the lubrication force begins to act. If particles are able to overcome the lubrication forces, 
a hertzian contact force is also added to ensure that there is no particle overlap, defined in Equation \eqref{eq:hertz}

@@

\begin{equation}
    \begin{split}
    \phi(\vec{r}_{ij}) &= {\epsilon} \tilde{\phi}\left(\frac{\vec{r}_{ij}}{{\sigma}}\right) , \\
    \vec{F}(\vec{r}_{ij}) &= {\epsilon} \tilde{\vec{F}}\left(\frac{\vec{r}_{ij}}{{\sigma}}\right) .
    \end{split}
\end{equation}

For the lubrication force \eqref{eq:lubrication}, we choose
${\sigma}=R_1+R_2$ and ${\epsilon}=\frac{6\pi\eta R_1^2 R_2^2}{{\sigma^3}}$, and for the
Hertz potential we chose ${\sigma}=R_1+R_2$ and ${\epsilon}=K_H\sigma^{5/2}$. For two identical, rotationally
symmetric ellipsoidal particles with orientations $\hat{\vec{o}}_i$ and $\hat{\vec{o}}_j$, we then replace $\epsilon$ and $\sigma$ by
the anisotropic functions

\begin{equation}
    \begin{split}
    \tilde\epsilon\left(\hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \frac{{\epsilon}}{\sqrt{1-\chi^2}} , \\
    %\qquad \chi = \frac{\left(\alpha^2-1\right)R_\parallel^2}{\left(\alpha^2+1\right)R_\parallel^2}\left(\hat{\vec{o}}_i\hat{\vec{o}}_j\right) , \\
    \tilde\sigma\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \frac{{\sigma}}{\sqrt{1-\frac{\chi}{2}\left[ \frac{\left(\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_i+\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_j\right)^2}{1+\chi\left(\hat{\vec{o}}_i\hat{\vec{o}}_j\right)} + \frac{\left(\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_i-\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_j\right)^2}{1-\chi\left(\hat{\vec{o}}_i\hat{\vec{o}}_j\right)} \right] }} , \\
    \chi &= \frac{\alpha^2-1}{\alpha^2+1} , \\
    \end{split}
\end{equation}

where $R_{\parallel}$ is the particle radius along the
symmetry axis $\hat{o}$ and $\alpha=\frac{R_{\parallel}}{R_{\perp}}$ the aspect
ratio of the particle. The anisotropic Hertz potential and lubrication
force are then defined as
%
\begin{equation}
    \begin{split}
    \phi\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \epsilon\left(\hat{\vec{o}}_i, \hat{\vec{o}}_j\right) \tilde{\phi}\left(\frac{\vec{r}_{ij}}{\sigma\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right)} \right) , \\
    \vec{F}\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \epsilon\left(\hat{\vec{o}}_i, \hat{\vec{o}}_j\right) \tilde{\vec{F}}\left(\frac{\vec{r}_{ij}}{\sigma\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right)} \right) .
    \end{split}
\end{equation}

@@

\begin{equation}
    U_{ij} = \frac{\mu_0 \vec{m}_i \vec{m}_j}{4\pi |r_{ij}|^{3}} \left[ \vec{o}_i \cdot \vec{o}_j - 
    3(\vec{o}_i \cdot \vec{r}_{ij})(\vec{o}_j \cdot \vec{r}_{ij}) \right]
    \label{eq:magnet_potential}
\end{equation}

Where $\mu_0 = 4\pi \cdot 10^{-7} \frac{H}{m}$,  $\Hat{\vec{o_i}}$ is the orientation unit vector of particle 
$i$, $\Hat{\vec{r}_{ij}}$ is the distance vector between particles $i$ and $j$ and $m_i$ is the magnitude the 
magnetic dipole of particle $i$. From the potential, the force and torque of the dipole force between particles 
can be found. These expressions are shown in equations \ref{eq:dipole_magnetic_force} and \ref{eq:dipole_magnetic_torque} 
for the force and torque respectively.

\begin{equation}
    \vec{F}_{ij} = \frac{3 \mu_0}{4 \pi} [\frac{5(\vec{m}_i \cdot \vec{r}_{ij} )(\vec{m}_j
    \cdot \vec{r}_{ij})}{|\vec{r}_{ij}|^7}\vec{r}_{ij} - \frac{(\vec{m}_{i} \cdot \vec{m}_{j})\vec{r}_{ij} + 
    (\vec{m}_{i} \cdot \vec{r}_{ij})\vec{m}_{i} + (\vec{m}_{j} \cdot \vec{r}_{ij})\vec{m}_{j} }{|\vec{r}_{ij}|^5}]
\label{eq:dipole_magnetic_force}
\end{equation}

\begin{equation}
    \vec{T}_{ij} = \frac{\mu_0}{4 \pi}[ \frac{3(\vec{m}_{j} \cdot \vec{r}_{ij})m_i \times \vec{r}_{ij} }
    {|\vec{r}_{ij}|^5} - \frac{\vec{m}_{i} \cdot \vec{m}_{j} }{|\vec{r}_{ij}|^3} ]
    \label{eq:dipole_magnetic_torque}
\end{equation}

@@

\begin{equation}
    \vec{F}_{j} = \vec{m}_{j}(\vec{o}_{j} \cdot \nabla \vec{B}_{i})
    \label{eq:magnet_force}
\end{equation}

\begin{equation}
    \vec{T}_{j} = \vec{m}_{j}(\vec{o}_{j} \times \vec{B}_{i})
    \label{eq:magnet_torque}
\end{equation}

@@

% \begin{table}[h!]
% \centering
% \begin{tabular}{||c c c c c c c c c c c c c||} 
%  \hline
%  $t_{sim}$ & $L_x$ & $\rho_f $ & $\tau$ & $\eta_f $ & $g_{cc'} \Delta m(\Delta t)^2/(\Delta x)^5$ & $\sigma$ & $\phi_f$ & $\rho_p$ & $\phi_p$ & $V_p (\Delta x)^3$ & $d_c \Delta x$ & $K_H$ \\ [0.5ex] 
%  \hline\hline
%  $10^5 \Delta t$ & 256$\Delta x$ & 0.7 $\frac{\Delta m}{(\Delta x)^3}$ & 1$\Delta t$ & 0.117 $\frac{\Delta m}{\Delta x \Delta t}$ & 0.08 & 0.0267 $\frac{\Delta m}{(\Delta t)^2}$ & 0.5 & 1 $\frac{\Delta m}{(\Delta x)^3}$ & 0.15 & 2000$\pi (\Delta x)^3$ & 2/3 $\Delta x$& 100 $(\Delta x)^{1/2}(\Delta t)^2/\Delta m$\\ [1ex] 
%  \hline
% \end{tabular}
% \caption{Summary of the fluid, particle and model parameters used in the proposed simulations}
% \label{table:model_params}
% \end{table}

\begin{table}
    \centering
    \caption{Summary of the parameters corresponding to the fluid, particle and model used in the simulations}
    \label{table:model_params}
    \begin{tabular}{|c|r|}
    \hline
    % Parameter & Value \\
    % \hline
    $L_x$ & 256 $\Delta x$ \\
    $\rho_f$ & 0.7 $ \frac{\Delta m}{(\Delta x)^3}$\\
    $\nu_f$ & 1/6 $\frac{\Delta x}{(\Delta t)^2}$ \\
    $\sigma $ & 0.0267 $\frac{\Delta m}{(\Delta t)^2}$ \\
    $\tau$ & 1 $\Delta t$\\
    $g_{kk'}$ & 0.08 $\frac{(\Delta x)^5}{\Delta m(\Delta t)^2}$\\
    $\phi_f$ & 0.5 \\
    $\phi_p$ & 0.15 \\
    $n_p$ & 1200 \\
    $\rho_p$ & 1 $ \frac{\Delta m}{(\Delta x)^3}$\\
    $V_p$ & $2000\pi/3 (\Delta x)^3$ \\
    $m$ & 1 $\frac{\Delta q}{\Delta t(\Delta x)^2}$\\
    $d_c$ & 2/3 $\Delta x$ \\
    $K_H$ & 100 $\frac{\Delta m}{(\Delta x)^{1/2}(\Delta t)^2} $\\
    \hline
    \end{tabular}
\end{table}

>>>>

\begin{equation}
    \sigma = -P\mathbb{I} + \eta\left( \nabla \vec{u} + (\nabla \vec{u})^T -\frac{2}{3}(\nabla \vec{u})\mathbb{I} \right)
\end{equation}

Substituting this stress tensor into the momentum equation and using the relation, $\nabla\cdot \vec{u} = 0$ from the incompressibility constraint, the incompressible 
Navier-Stokes equation can be derived given by,

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = -\nabla P + \eta \nabla^2 \vec{u} + \vec{f}
\end{equation}

To facilitate numerical simulation and analysis, it is often useful to express the Navier-Stokes equation in non-dimensional form. Introducing characteristic scales for length 
$l_0$, velocity $u_0$, and density $\rho_0$, dimensionless variables are defined as

\begin{equation}
    \bar{l} = \frac{l}{l_0}, \quad \bar{u} = \frac{u}{u_0}, \quad \bar{t} = \frac{u_0 t}{l_0} \quad \bar{\rho} = \frac{\rho}{\rho_0}, \quad \bar{P} = \frac{P}{\rho u^2}, 
\end{equation}

The form of the dimensionless pressure $\bar{P}$ is not unique and depends on whether inertial or viscous effects dominate. The expression shown is 
relevant to an inertial regime. Substituting these into the incompressible Navier-Stokes equation and approximating that no external forces are present 
leads to the dimensionless form:

\begin{equation}
    \frac{\partial \bar{\vec{u}}}{\partial \bar{t}} + (\bar{\vec{u}} \cdot\nabla)\bar{\vec{u}} = \frac{1}{Re} \nabla^2 \bar{\vec{u}} - \nabla\bar{P}
\end{equation}

where the Reynolds number is defined as $Re = \frac{\rho_0 u_0 l_0}{\eta}$ which characterizes the ratio of inertial to viscous forces in the system. For a system
where viscous effects dominate, $Re << 1$ the dimensionless pressure is defined as $\bar{P} = \frac{P l_0}{\eta u_0}$. This yields the alterate form,

\begin{equation}
    Re\left( \frac{\partial \bar{\vec{u}}}{\partial \bar{t}} + (\bar{\vec{u}} \cdot\nabla)\bar{\vec{u}}  \right) = \nabla^2 \bar{\vec{u}} - \nabla\bar{P}
\end{equation}

When substituting the value of $Re$ into the above, the Stokes equation is recovered.

@@

Bijels form through spinodal decomposition of two partially miscible fluids, a process that can be effectively modeled using the Cahn-Hilliard equation. \cite{cahn_spinodal_1961} While the
original equation is written for a concentration of species $c$. For a binary mixture, it can be recast in terms of a conserved scalar order parameter $\phi$, typically representing the 
local concentration difference between the two fluid components. As we are also accounting for the effect of hydrodynamics on the flow of mass, the advection term
from the mass flux $\vec{J}$ cannot be set to 0. Therefore the advective Cahn-Hilliard equation of an incompressible fluid ($\nabla \cdot \vec{v} = 0$) is given by,

\begin{equation}
    \begin{split}
        \nabla \cdot \vec{J} &= \nabla \cdot \phi\vec{u} - M \nabla \mu \\
        \frac{\partial \phi}{\partial t} + \nabla \cdot \vec{J} &= 0 \\
        \frac{\partial \phi}{\partial t} + \vec{u}\nabla \cdot \phi &= \nabla \cdot \left( M \nabla \mu \right)
    \end{split}
\end{equation}

Here, $M$ is the mobility coefficient and $\mu$ is the chemical potential derived from the variation of the free energy functional $F[\phi]$ which is the sum of a bulk free energy
$f(\phi)$ and an interfacial contribution. This formulation allows for the modeling of interfacial tension and the spontaneous formation of bicontinuous domains characteristic of spinodal decomposition.
To capture not only the diffusive separation of phases but also the resulting fluid motion, the Cahn-Hilliard equation is coupled with the Navier-Stokes equation. This is done
through the pressure which is expressed as the sum of the mechanical $P$ and chemical pressure $P_{chem}$ contributions to yield $P_{tot} = P + P_{chem}$. For a binary system, $P_{chem}$ 
can be written as a function of $\phi$ given by,

\begin{equation}
    P_{chem} = \left[ \phi \mu - f(\phi) - \kappa(\phi \nabla^2 \phi + \frac{1}{2}|\nabla \phi |^2 )\right] \mathbb{I} + \kappa \nabla \phi \otimes \nabla \phi
\end{equation}

The divergence of the above expression is found to be $\nabla \cdot P_{chem} = \phi \nabla \mu$. Substituting $P_{tot}$ for $P$ in the Navier Stokes equation,
the Cahn-Hilliard-Navier-Stokes (CHNS) framework is obtained. This coupling is essential in systems like bijels, where advection plays a role in domain 
coarsening. The momentum conservation equation in this framework is given by,

\begin{equation}
    \rho \left(\frac{\partial\vec{u}}{\partial t} + (\vec{u}\cdot\nabla)\vec{u} \right) = \eta \nabla^2 \vec{u} - \nabla P - \phi \nabla \mu
\end{equation}

@@

However the BGK operator controls all moments at a single relaxation rate which can induce numerical instabilities and inaccuracies
at higher fluid velocities. \cite{liu_simulation_2023, adhikari_fluctuating_2005} These numerical instabilities can be resolved by using
a Multiple Relaxation Time (MRT) collision operator. However the relaxation rates used in an MRT collision operator must be tuned carefully
to ensure that the appropriate fluid properties are modelled. Simpler variants of the MRT collision operator such as the Two Relaxation Time (TRT)
collision operator have been proposed as a solution that provides some of the benefits of MRT models with the simplicity of SRT models. These 
have facilitated simulation of a wider range of physical phenomena \cite{adhikari_fluctuating_2005, liu_simulation_2023}. The regimes that the simulations 
performed here will take place will be at relatively small Reynolds numbers, mitigating many of the weaknesses of the SRT model while being computatinally
more efficient. Therefore, the SRT collision operator is selected. The combined streaming and 
collision step in LBM, using the BGK operator, is expressed in Equation \ref{eq:LBM_BGK}.

@@

where $w_i$ and $\vec{c}_i$ correspond to the weights and lattice velocity vector of each direction in the D3Q19 lattice shown in Figure \ref{fig:d3q19_lattice}, $\rho$ and $\vec{u}$ are the fluid density
and velocity at each lattice point, and $c_s$ is the speed of sound of the lattice, set to $c_s = \frac{1}{\sqrt{3}}$. $\rho$ and $\vec{u}$ can be calculated from the zeroth and first moment of the 
mass distribution $f_i$, given by

@@

Using the Chapman-Enskog expansion, the Navier-Stokes equation at the incompressible limit at low Mach
numbers can be recovered. \cite{qian_lattice_1992, he_lattice_1997} The regime accessible through this method is suitable 
for simulations in this work as the $ 1 \leq Re \leq 100 $ and the mach number, $Ma < 0.01$, fulfilling the stability 
criterion and usage requirements for the presented hydrodynamic model.
The momentum flux tensor is calculated from the Chapman-Enskog expansion of \ref{eq:LBM_BGK} to the second order yielding,

@@

The momentum flux tensor can be expressed as a sum of the equilibrium and non-equilibrium components of the stress tensor, denoted by 0 and 1 respectively given by,

@@

An effective mass of each fluid at node $\vec{x}$ is used in place of the density to scale it between zero 
and one and is defined as $\psi^{k}(\vec{x},t) = \rho_{0}\left[1 - \exp(-\frac{\rho^{k}(\vec{x}, t)}{\rho_{0}})\right]$ where
$\rho_0$ is the reference density and is set to $\rho_0 = 1 \frac{\Delta m}{\Delta x^3}$. 
In this model, the SC force is incorporated into the macroscopic velocities that are then used to calculate the equilibrium
distribution $f_{i}^{k, eq}$ for fluid $k$, defined as,

@@

The particles are discretized on the lattice according to the method laid out in Ladd 1994 and Aidun 1998
\cite{ladd_numerical_1994,aidun_direct_1998}. Nodes representing the particle are marked as solid nodes that replicate
a no-slip boundary condition through a moving bounce-back boundary. This is implemented into the distribution function
by reflecting the outgoing populations of $f_i^k$ to the opposite lattice velocity $f^k_{i^\star}$ given by

@@

where $\vec{u}_i = \vec{u}_p + \vec{omega}\times \vec{r_i}$ which represents the difference between the particle velocity and the 
velocity of the particle surface. The location of the particle surface $r_i$ is calculated as $r_i = \vec{x} + \frac{\Delta t}{2}\vec{c}_i - \vec{x}_{cm}$ 
where $\vec{x}_{cm}$ is the center of mass of the particle.
This facilitates momentum exchange between particle and fluid which can be calculated analytically as 
\(\Delta\vec{p}^k_i = 2 f^{k,\star}_i(\vec{x},t)\vec{c}_i - \frac{2w_i}{c_s^2}\rho(\vec{u}_i\cdot\vec{c}_i)\vec{c}_i\).
The sum of the momentum change across the surface of the particle is computed to obtain the force and torque on the particle,

@@

where $N_{\text{f}}$ is the number of fluid nodes around the node being uncovered and $\vec{c}_{i_{\text{f}}n}$ is the direction of the lattice velocity 
in the grid point being uncovered that points to a surrounding fluid node.

@@

where $R_i$ and $R_j$ are the radii of each particle involved in the interaction, $\vec{r}_{ij}$ is the distance
vector between the particle centers, $\vec{u}_{ij}$ are the relative velocities of the particles and $d_c$ 
is the cutoff distance when the lubrication force begins to act. If particles are able to overcome the lubrication forces, 
a hertzian contact force is also added to ensure that there is no particle overlap, defined in Equation \eqref{eq:hertz}

@@

\begin{equation}
    \begin{split}
    \phi(\vec{r}_{ij}) &= {\epsilon} \tilde{\phi}\left(\frac{\vec{r}_{ij}}{{\gamma}}\right) , \\
    \vec{F}(\vec{r}_{ij}) &= {\epsilon} \tilde{\vec{F}}\left(\frac{\vec{r}_{ij}}{{\gamma}}\right) .
    \end{split}
\end{equation}

For the lubrication force \eqref{eq:lubrication}, we choose
${\gamma}=R_1+R_2$ and ${\epsilon}=\frac{6\pi\eta R_1^2 R_2^2}{{\gamma^3}}$, and for the
Hertz potential we chose ${\gamma}=R_1+R_2$ and ${\epsilon}=K_H\gamma^{5/2}$. For two identical, rotationally
symmetric ellipsoidal particles with unit orientation vectors $\hat{\vec{o}}_i$ and $\hat{\vec{o}}_j$, we then replace $\epsilon$ and $\gamma$ by
the anisotropic functions

\begin{equation}
    \begin{split}
    \tilde\epsilon\left(\hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \frac{{\epsilon}}{\sqrt{1-\chi^2}} , \\
    %\qquad \chi = \frac{\left(\alpha^2-1\right)R_\parallel^2}{\left(\alpha^2+1\right)R_\parallel^2}\left(\hat{\vec{o}}_i\hat{\vec{o}}_j\right) , \\
    \tilde\gamma\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \frac{{\gamma}}{\sqrt{1-\frac{\chi}{2}\left[ \frac{\left(\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_i+\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_j\right)^2}{1+\chi\left(\hat{\vec{o}}_i\hat{\vec{o}}_j\right)} + \frac{\left(\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_i-\hat{\vec{r}}_{ij}\cdot\hat{\vec{o}}_j\right)^2}{1-\chi\left(\hat{\vec{o}}_i\hat{\vec{o}}_j\right)} \right] }} , \\
    \chi &= \frac{\alpha^2-1}{\alpha^2+1} , \\
    \end{split}
\end{equation}

where $R_{\parallel}$ is the particle radius along the
symmetry axis and $\alpha=\frac{R_{\parallel}}{R_{\perp}}$ the aspect
ratio of the particle. The anisotropic Hertz potential and lubrication
force are then defined as
%
\begin{equation}
    \begin{split}
    \phi\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \epsilon\left(\hat{\vec{o}}_i, \hat{\vec{o}}_j\right) \tilde{\phi}\left(\frac{\vec{r}_{ij}}{\gamma\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right)} \right) , \\
    \vec{F}\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right) &= \epsilon\left(\hat{\vec{o}}_i, \hat{\vec{o}}_j\right) \tilde{\vec{F}}\left(\frac{\vec{r}_{ij}}{\gamma\left(\vec{r}_{ij}, \hat{\vec{o}}_i, \hat{\vec{o}}_j\right)} \right) .
    \end{split}
\end{equation}

@@

\begin{equation}
    U_{ij} = \frac{\mu_0 m_i m_j}{4\pi |r_{ij}|^{3}} \left[ \Hat{\vec{o_i}}\cdot \Hat{\vec{o_j}} - 
    3(\Hat{\vec{o_i}} \cdot \Hat{\vec{r_{ij}}})(\Hat{\vec{o_j}} \cdot \Hat{\vec{r_{ij}}}) \right]
    \label{eq:magnet_potential}
\end{equation}

Where $\mu_0 = 4\pi \cdot 10^{-7} \frac{H}{m}$,  $\Hat{\vec{o_i}}$ is the orientation unit vector of particle 
$i$, $\Hat{\vec{r_{ij}}}$ is the unit distance vector between particles $i$ and $j$ and $m_i$ is the magnitude the 
magnetic dipole of particle $i$. From the potential, the force and torque of the dipole force between particles 
can be found. These expressions are shown in equations \ref{eq:dipole_magnetic_force} and \ref{eq:dipole_magnetic_torque} 
for the force and torque respectively.

\begin{equation}
    \vec{F}_{ij} = \frac{3 \mu_0 m_i m_j}{4 \pi} [\frac{5( \Hat{\vec{o_i}} \cdot \Hat{\vec{r_{ij}}} )(\Hat{\vec{o_j}}
    \cdot \Hat{\vec{r_{ij}}})}{|\vec{r_{ij}}^4|}\Hat{\vec{r_{ij}}} - \frac{(\Hat{\vec{o_i}} \cdot \Hat{\vec{o_j}})\Hat{\vec{r_{ij}}} + 
    (\Hat{\vec{o_i}} \cdot \Hat{\vec{r_{ij}}})\Hat{\vec{o_j}} + (\Hat{\vec{o_j}} \cdot \Hat{\vec{r_{ij}}})\Hat{\vec{o_j}} }{|\vec{r_{ij}}|^4}]
\label{eq:dipole_magnetic_force}
\end{equation}

\begin{equation}
    \vec{T}_{ij} = \frac{\mu_0 m_i m_j}{4 \pi}[ \frac{3(\Hat{\vec{o_j}} \cdot \Hat{\vec{r_{ij}}})\Hat{\vec{o_i}} \times \Hat{\vec{r_{ij}}} }
    {|\vec{r}_{ij}|^3} - \frac{\Hat{\vec{o_i}} \cdot \Hat{\vec{o_j}} }{|\vec{r}_{ij}|^3} ]
    \label{eq:dipole_magnetic_torque}
\end{equation}

@@

\begin{equation}
    \vec{F}_{j} = m_j(\Hat{\vec{o_j}} \cdot \nabla \vec{B}_{i})
    \label{eq:magnet_force}
\end{equation}

\begin{equation}
    \vec{T}_{j} = m_j(\Hat{\vec{o_j}} \times \vec{B}_{i})
    \label{eq:magnet_torque}
\end{equation}

@@

\begin{table}
    \centering
    \caption{Summary of the parameters corresponding to the fluid, particle and model used in the simulations. 
             LB parameters can be converted to their real world equivalents using the unit conversions provided.}
    \label{table:model_params}
    \begin{tabular}{|c|c|}
    \hline
    Parameter & Value \\ [0.5ex] 
    \hline\hline
    $L_x$ & 256 $\Delta x$ \\
    $\rho_f$ & 0.7 $ \frac{\Delta m}{(\Delta x)^3}$\\
    $\nu_f$ & 1/6 $\frac{\Delta x}{(\Delta t)^2}$ \\
    $\sigma $ & 0.0267 $\frac{\Delta m}{(\Delta t)^2}$ \\
    $\tau$ & 1 $\Delta t$\\
    $g_{kk'}$ & 0.08 $\frac{(\Delta x)^5}{\Delta m(\Delta t)^2}$\\
    $\phi_f$ & 0.5 \\
    $\phi_p$ & 0.15 \\
    $n_p$ & 1200 \\
    $\rho_p$ & 1 $ \frac{\Delta m}{(\Delta x)^3}$\\
    $V_p$ & $2000\pi/3 (\Delta x)^3$ \\
    $m$ & 1 $\frac{\Delta q}{\Delta t(\Delta x)^2}$\\
    $d_c$ & 2/3 $\Delta x$ \\
    $K_H$ & 100 $\frac{\Delta m}{(\Delta x)^{1/2}(\Delta t)^2} $\\
    \hline
    \end{tabular}
\end{table}

In these simulations, the aspect ratio is defined as $\alpha = \frac{R_{\parallel}}{R_{\perp}}$ with $\alpha = 0.5, 1, 2$ 
used in these simulations. All particle geometries are volume matched, based on the spherical particle volume. For the 
spherical particle, the radius was set at $R_{\parallel} = R_{\perp} = 7.9 \Delta x$ which sets the $\alpha = 0.5$ particles to 
$R_{\parallel} = 5 \Delta x, R_{\perp} = 10 \Delta x$ and the $\alpha = 2$ particles to $R_{\parallel} = 12.6 \Delta x, R_{\perp} = 6.3 \Delta x$. 
$R_{\parallel}$ and $R_{\perp}$ is the radius of the particle parallel and perpendicular to the symmetry axis of 
the particle respectively. These values are identical to what was used in Gunther et al, allowing comparisons to the 
trends they observed. \cite{gunther_timescales_2014} The particles were randomly placed in the system and an equilibration
step was performed to push the particles further away from each after placement. Once equilibration is performed, a 
constant magnetic field is switched on with dimensionless strengths defined using $\bar{B}$.
